### 2017-04-16 paper/oopsla17-submission/conclusion.tex
asd

We have presented \toolname, which
combines modern statistical methods
with domain-specific feature engineering
to open the door to a new data-driven
path to precise error localization.
%
\toolname
translates pairs of
blame-labeled ill-typed
programs into bags-of-abstracted-terms,
and then trains a classifier
to predict whether a feature vector
should get a blame label.
% that
% predicts the value of the
% whether a feature vector gets
% a blame label.
%
Given a new ill-typed program,
we extract its bag-of-abstracted-terms
to compute the blame label for each
term and rank the predictions based
on the model's confidence, to select
the top 1--3 blame assignments to
present to the user.

We have evaluated our approach with
five different classifiers on a suite
of over 4,500 student programs.
%
We found that while machine learning
over syntactic features of each term in isolation
performs worse than existing
purely constraint based approaches, % (\eg \ocaml, \sherrloc),
augmenting the data with a single feature corresponding to
the type error slice brings our
classifiers up to par with the state of the art,
and further augmenting the data with
features of an expression's parent and children
allows our classifiers to outperform
the state of the art by 18 percentage points.

@@

As with other forms of machine learning,
a key concern is that of \emph{bias}: are
\toolname's models specific
to our data set, would they fail on
\emph{other} programs?
%
We address this concern in two ways.
%
First, we partition the data by year,
and show that models learned from one
year generalize to, \ie perform nearly
as well on, the programs from the other
year.
%
Second, we argue that in our setting
this bias is a \emph{feature} (and not
a bug): it allows \toolname to \emph{adapt}
to the kinds of errors that programmers
(specifically novices, who are in greatest
need of precise feedback) actually make,
rather than hardwiring the biases of
experts who % , by dint of their
% training and experience,
may suffer from blind spots. % with regards to such problems.
%
In this regard, we are particularly pleased
that our classifiers can be trained on a
modest amount of data, \ie a single course's
worth, and envision a future where each course
comes equipped with a model of its students' errors.

>>>>

We have presented \toolname, which
combines modern statistical methods
with domain-specific feature engineering
to open the door to a new data-driven
path to precise error localization,
that significantly outperforms the
state of the art on a new benchmark
suite comprising 4500 student programs.
%
%%% RJNUKEDME   \toolname
%%% RJNUKEDME   translates pairs of
%%% RJNUKEDME   blame-labeled ill-typed
%%% RJNUKEDME   programs into bags-of-abstracted-terms,
%%% RJNUKEDME   and then trains a classifier
%%% RJNUKEDME   to predict whether a feature vector
%%% RJNUKEDME   should get a blame label.
%%% RJNUKEDME   % that
%%% RJNUKEDME   % predicts the value of the
%%% RJNUKEDME   % whether a feature vector gets
%%% RJNUKEDME   % a blame label.
%%% RJNUKEDME   %
%%% RJNUKEDME   Given a new ill-typed program,
%%% RJNUKEDME   we extract its bag-of-abstracted-terms
%%% RJNUKEDME   to compute the blame label for each
%%% RJNUKEDME   term and rank the predictions based
%%% RJNUKEDME   on the model's confidence, to select
%%% RJNUKEDME   the top 1--3 blame assignments to
%%% RJNUKEDME   present to the user.
%%% RJNUKEDME
%%% RJNUKEDME   We have evaluated our approach with
%%% RJNUKEDME   five different classifiers on a suite
%%% RJNUKEDME   of over 4,500 student programs.
%%% RJNUKEDME   %
%%% RJNUKEDME   We found that while machine learning
%%% RJNUKEDME   over syntactic features of each term in isolation
%%% RJNUKEDME   performs worse than existing
%%% RJNUKEDME   purely constraint based approaches, % (\eg \ocaml, \sherrloc),
%%% RJNUKEDME   augmenting the data with a single feature corresponding to
%%% RJNUKEDME   the type error slice brings our
%%% RJNUKEDME   classifiers up to par with the state of the art,
%%% RJNUKEDME   and further augmenting the data with
%%% RJNUKEDME   features of an expression's parent and children
%%% RJNUKEDME   allows our classifiers to outperform
%%% RJNUKEDME   the state of the art by 18 percentage points.

@@

As with other forms of machine learning,
a key concern is that of \emph{data-set bias}: are
\toolname's models specific
to our data set, would they fail on
\emph{other} programs?
%
We address this concern in two ways.
%
First, we partition the data by year,
and show that models learned from one
year generalize to, \ie perform nearly
as well on, the programs from the other
year.
%
Second, we argue that in our setting
this bias is a \emph{feature} (and not
a bug): it allows \toolname to \emph{adapt}
to the kinds of errors that programmers
(specifically novices, who are in greatest
need of precise feedback) actually make,
rather than hardwiring the biases of
experts who % , by dint of their
% training and experience,
may suffer from blind spots. % with regards to such problems.
%
In this regard, we are particularly pleased
that our classifiers can be trained on a
modest amount of data, \ie a single course's
worth, and envision a future where each course
comes equipped with a model of its students' errors.

### 2017-04-16 paper/oopsla17-submission/evaluation.tex
asd

\section{Quantitative Evaluation}
\label{sec:evaluation}
\input{data}

We have implemented our technique for localizing type errors for a
purely functional subset of \ocaml with polymorphic types and functions.
%
We seek to answer three questions in our evaluation:
%
\begin{description}
\item[1. Blame Accuracy (\autoref{sec:quantitative})]
  %
  How often do we \emph{correctly} assign blame for the type error?
  %
  % We compare our technique with a variety of off-the-shelf classifiers
  % and find that our top-ranked blame assignments have an accuracy of
  % 72\%, compared to a state-of-the-art 56\%.
  % For how many ill-typed programs can we accurately predict the source
  % of the error?
\item[2. Feature Utility (\autoref{sec:feature-utility})]
  %
  Which feature sets are \emph{most important} for accurate predictions?
  % How much do the features described in \autoref{sec:features}
  % contribute to our predictions?
\item[3. Explaining Predictions (\autoref{sec:qualitative})]
  %
  Can we \emph{explain} individual predictions (correct or not)?
\end{description}
%\ES{may want to swap (1) and (2)..}

>>>>

\mysection{Quantitative Evaluation}
\label{sec:evaluation}
\input{data}

We have implemented our technique for localizing type errors for a
purely functional subset of \ocaml with polymorphic types and functions.
%
We seek to answer three questions in our evaluation:
%
\begin{itemize}
\item \textbf{Blame Accuracy}
  %
  How often does \toolname
  blame a \emph{correct}
  location for the type error?
  (\autoref{sec:quantitative})
  %
  % We compare our technique with a variety of off-the-shelf classifiers
  % and find that our top-ranked blame assignments have an accuracy of
  % 72\%, compared to a state-of-the-art 56\%.
  % For how many ill-typed programs can we accurately predict the source
  % of the error?
\item \textbf{Feature Utility}
  %
  Which program \emph{features are required}
  to localize errors?
   (\autoref{sec:feature-utility})
  % How much do the features described in \autoref{sec:features}
  % contribute to our predictions?
\item \textbf{Interpretability}
  %
  %% Do the models learned by \toolname
  %% correspond to our intuition about
  %% the real causes of errors?
  Are the models \emph{interpretable} using
  our intuition about the causes of type errors?
  (\autoref{sec:qualitative})
\end{itemize}
%
%\mypara{Summary of Results}
%
In the sequel we present our experimental
methodology \autoref{sec:methodology} and
then drill into how we evaluated each of
the questions above.
%
However, for the impatient reader, we begin
with a quick summary of our main results:
%
%
%\begin{itemize}
%
%\item \textbf

\mypara{1. Data beats Algorithms}
Our main result is that for type error
localization, data is indeed unreasonably
effective~\citep{halevy09}.
%
\toolname's most sophisticated
\emph{neural network}
based classifier's top-1 error
location blames the correct
sub-term 74\% of the time
which, a good 18 points
higher than the state of
the art \sherrloc's 56\%.
%
However, even \toolname's simple
\emph{logistic regression} based
classifier is correct 60\%,
\ie 4 points better than \sherrloc.
%
When the top-3 locations are considered,
\toolname is correct 91\% of the time.

% \item \textbf
\mypara{2. Slicing is Critical}
%
However, data is effective \emph{only}
when irrelevant sub-terms have been
sliced out of consideration.
%
In fact, perhaps our most surprising
result is that type-error-slicing and
local syntax alone yields
a neural classifier that is 10 points
better than \ocaml and on par with
\sherrloc.
%
That is once we focus classifiers on
slices, purely syntactic features
alone perform as well as the
state-of-the-art.

%\item \textbf
\mypara{3. Size doesn't matter, types do}
%
We find that (after slices)
contextual typing features
provide the biggest
improvement in accuracy.
%
Furthermore we find contextual syntax
features to be relatively unimportant,
which would support the hypothesis that
the context's \emph{type} nicely
summarizes the properties of the
surrounding expressions.
%
Finally, we found that the \emph{size}
of the sub-expression was not a useful
feature. This was unexpected as we thought
smaller expressions would be simpler and
hence, more likely causes.

% \item \textbf
\mypara{4. Models Learn Typing Rules}
%
Finally, by investigating a few of the
predictions made by the \emph{decision tree}
based models, we found that the models
do indeed capture some simple and intuitive
rules for predicting well-typededness.
%
For example, in an application, if the
left argument is a function then the
error is likely on the right sub-term;
in a function definition \RJ{fixme: orig
defining a function is a fine thing
to do?}
%
%\end{itemize}

\RJ{eric please check the numbers -- we say 71, 72, 74, 91, 92, 94?}

### 2017-04-16 paper/oopsla17-submission/related.tex
asd

In this section we describe two relevant aspects of related work:
%
programming languages approaches to diagnosing type errors, and
%
software engineering approaches to fault localization.
%
% applying machine learning techniques to problems in programming languages.
%
% \mysubsection{Type Error Diagnosis}
Languages with static type systems and type inference produce type
errors that novices often perceive as difficult to
interpret~\citep{Wand1986-nw}.
%
For example, approaches based on Hindley-Milner or other constraint
systems typically have the issue that when the constraint $\alpha=\beta$
is violated, the system cannot know whether the user should intend to
fix $\alpha$ or $\beta$ and must thus report the discrepancy in a
generic manner.
%
As a result, a number of approaches have been proposed to
more precisely localize the type error,
give clearer error messages, or
fix the error automatically.
%
% The technique most related to our work is \ES{TODO}.  It has been shown
% to be quite good at \ES{TODO}. However, we choose to focus on \ES{TODO},
% and our approach contains \ES{TODO}, which is not present in previous
% work.

\mypara{Localizing Type Errors}
It is well-known that the original Damas-Milner algorithm $\mathcal{W}$
produces poor error locations~\citep{Wand1986-nw}, often reported far
from their source.
%
The type checker collects typing constraints as it traverses the
program, and it crucially assumes that if a constraint can be safely
added to the set of assumptions (\ie no type error has been found yet),
then the constraint can be \emph{assumed} to be correct.
%
The type checker reports an error the moment it finds a constraint that
contradicts one of the assumptions, blaming the new inconsistent
constraint, and thus it is extremely sensitive to the order in which it
traverses the source program (the infamous ``left-to-right''
bias~\citep{McAdam1998-ub}).
%
Several alternative traversal have been proposed, \eg
top-down rather than bottom-up~\citep{Lee1998-ys}, or a \emph{symmetric}
traversal that checks sub-expressions independently and only reports an
error when two inconsistent sets of constraints are
merged~\citep{McAdam1998-ub,Yang1999-yr}.

\mypara{Slicing}
Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
overcomes this bias by extracting a complete and minimal subset
of terms that contribute to the error, \ie all of the
terms that are required for it to manifest and no more.
%
Slicing typically requires rewriting the type checker with a
specialized constraint language and solver, though
\citet{Schilling2011-yf} shows how to turn any type checker into a
slicer by treating it as a black-box.
%
While slicing techniques guarantee enough information to diagnose the
error, they can fall into the trap of providing \emph{too much}
information, producing a slice that is not much smaller than
the input. %original. % input program.

@@

\mypara{Explaining Type Errors}
In this paper we have focused solely on the task of \emph{localizing} a
type error, but a good error report should also \emph{explain} the
error.
%
\citet{Wand1986-nw}, \citet{Beaven1993-hb}, and \citet{Duggan1996-by}
attempt to explain type errors by collecting the chain of inferences
made by the type checker %--- essentially the typing derivation ---
and presenting them to the user.
%
% Such explanations can be lengthy, as an attempt to compress the
% explanation \citet{Yang2000-kz} presents a visualization of the
% inference process.
%
\citet{Gast2004-zd} produces a slice enhanced by arrows
showing the dataflow from sources with different types to a
shared sink, borrowing the insight of dataflows-as-explanations from
\textsc{MrSpidey}~\citep{Flanagan1996-bu}.

\mypara{Interactive Explanations}
Such \emph{static} explanations of type errors run the risk of
overwhelming the user with too much information, it may be preferable to
treat type error diagnosis as an \emph{interactive} debugging session.
%
\citet{Bernstein1995-yj} extend the type inference procedure to handle
\emph{open} expressions (\ie with unbound variables), allowing users to
interactively query the type checker for the types of sub-expressions.
%
\citet{Chitil2001-td} proposes \emph{algorithmic debugging} of type
errors, presenting the user with a sequence of yes-or-no questions about
the inferred types of sub-expressions that guide the user to a specific
explanation for the error.
%
\citet{Seidel2016-ul} explain type errors by searching for inputs that
expose the \emph{run-time} error that the type system prevented, and
present users with an interactive visualization of the erroneous
computation.

\mypara{Programmatic Explanations}
%
The best explanation of a type error, however, might be given by an
expert, \eg the compiler or library author.
%
\citet{Hage2006-hc} catalog a set of heuristics for
improving the quality of error messages by examining errors made by
novices.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} extend the ability to customize error messages to
library authors, enabling \emph{domain-specific} errors.
%
% The 8.0 release of the
% Glasgow Haskell Compiler\footnote{\url{https://ghc.haskell.org/trac/ghc/wiki/Proposal/CustomTypeErrors}}
% incorporates these ideas, allowing library authors to supply
% custom errors when type-class resolution or type-family reduction fail,
% but not for ordinary unification failures.

>>>>

In this section we describe two relevant aspects of related work:
%
programming languages approaches to diagnosing type errors, and
%
software engineering approaches to fault localization.
%
% applying machine learning techniques to problems in programming languages.
%
% \mysubsection{Type Error Diagnosis}
%% Languages with static type systems and type inference produce type
%% errors that novices often perceive as difficult to
%% interpret~\citep{Wand1986-nw}.
%% %
%% For example, approaches based on Hindley-Milner or other constraint
%% systems typically have the issue that when the constraint $\alpha=\beta$
%% is violated, the system cannot know whether the user should intend to
%% fix $\alpha$ or $\beta$ and must thus report the discrepancy in a
%% generic manner.
%% %
%% As a result, a number of approaches have been proposed to
%% more precisely localize the type error,
%% give clearer error messages, or
%% fix the error automatically.
%
% The technique most related to our work is \ES{TODO}.  It has been shown
% to be quite good at \ES{TODO}. However, we choose to focus on \ES{TODO},
% and our approach contains \ES{TODO}, which is not present in previous
% work.

\mypara{Localizing Type Errors}
It is well-known that the original Damas-Milner
algorithm $\mathcal{W}$ produces errors far
from their source, that novices percieve as
difficult to interpret~\citep{Wand1986-nw}.
%
%% The type checker collects typing constraints as it traverses the
%% program, and it crucially assumes that if a constraint can be safely
%% added to the set of assumptions (\ie no type error has been found yet),
%% then the constraint can be \emph{assumed} to be correct.
%
The type checker reports an error the moment
it finds a constraint that contradicts one
of the assumptions, blaming the new inconsistent
constraint, and thus it is extremely sensitive
to the order in which it traverses the source
program (the infamous ``left-to-right''
bias~\citep{McAdam1998-ub}).
%
Several alternative traversal have been proposed,
\eg top-down rather than bottom-up~\citep{Lee1998-ys},
or a \emph{symmetric} traversal that checks
sub-expressions independently and only reports an
error when two inconsistent sets of constraints are
merged~\citep{McAdam1998-ub,Yang1999-yr}.
%
%\mypara{Slicing}
Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
overcomes the constraint-order bias by extracting a
complete and minimal subset
of terms that contribute to the error, \ie all of the
terms that are required for it to manifest and no more.
%
Slicing typically requires rewriting the type checker with a
specialized constraint language and solver, though
\citet{Schilling2011-yf} shows how to turn any type checker into a
slicer by treating it as a black-box.
%
While slicing techniques guarantee enough information to diagnose the
error, they can fall into the trap of providing \emph{too much}
information, producing a slice that is not much smaller than
the input. %original. % input program.

@@

\mypara{Explaining Type Errors}
In this paper we have focused solely on the task of \emph{localizing} a
type error, but a good error report should also \emph{explain} the
error.
%
\citet{Wand1986-nw}, \citet{Beaven1993-hb}, and \citet{Duggan1996-by}
attempt to explain type errors by collecting the chain of inferences
made by the type checker %--- essentially the typing derivation ---
and presenting them to the user.
%
% Such explanations can be lengthy, as an attempt to compress the
% explanation \citet{Yang2000-kz} presents a visualization of the
% inference process.
%
\citet{Gast2004-zd} produces a slice enhanced by arrows
showing the dataflow from sources with different types to a
shared sink, borrowing the insight of dataflows-as-explanations from
\textsc{MrSpidey}~\citep{Flanagan1996-bu}.
%
\citet{Hage2006-hc} catalog a set of heuristics for
improving the quality of error messages by examining errors made by
novices.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} extend the ability to customize error messages to
library authors, enabling \emph{domain-specific} errors.
%
%\mypara{Interactive Explanations}
Such \emph{static} explanations of type errors run the risk of
overwhelming the user with too much information, it may be preferable to
treat type error diagnosis as an \emph{interactive} debugging session.
%
\citet{Bernstein1995-yj} extend the type inference procedure to handle
\emph{open} expressions (\ie with unbound variables), allowing users to
interactively query the type checker for the types of sub-expressions.
%
\citet{Chitil2001-td} proposes \emph{algorithmic debugging} of type
errors, presenting the user with a sequence of yes-or-no questions about
the inferred types of sub-expressions that guide the user to a specific
explanation for the error.
%
\citet{Seidel2016-ul} explain type errors by searching for inputs that
expose the \emph{run-time} error that the type system prevented, and
present users with an interactive visualization of the erroneous
computation.

%% RJ-CUT-ME \mypara{Programmatic Explanations}
%% RJ-CUT-ME %
%% RJ-CUT-ME The best explanation of a type error, however, might be given by an
%% RJ-CUT-ME expert, \eg the compiler or library author.
%% RJ-CUT-ME %
%% RJ-CUT-ME \citet{Hage2006-hc} catalog a set of heuristics for
%% RJ-CUT-ME improving the quality of error messages by examining errors made by
%% RJ-CUT-ME novices.
%% RJ-CUT-ME %
%% RJ-CUT-ME \citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
%% RJ-CUT-ME \citet{Serrano2016-oo} extend the ability to customize error messages to
%% RJ-CUT-ME library authors, enabling \emph{domain-specific} errors.
%
% The 8.0 release of the
% Glasgow Haskell Compiler\footnote{\url{https://ghc.haskell.org/trac/ghc/wiki/Proposal/CustomTypeErrors}}
% incorporates these ideas, allowing library authors to supply
% custom errors when type-class resolution or type-family reduction fail,
% but not for ordinary unification failures.
